Statically Typed String Sanitation Inside a Python

### 2014-08-19 sanitation-psp14/sanitation-psp14-short.tex
More layout work

Figure \ref{examplefig} demonstrates the use of string types in atlang.
The sanitize function takes an arbitrary string and returns a string without
double quotes or left and right brackets. In this example, we use HTML
escape sequences.

@@

The result of applying sanitize to input is appended to two functions which
construct a safe query and safe output.
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
XSS attacks.

Note that ${\sf input}$ does not actually meet these specifications without additional machinery.
The type of ${\sf input}$ is quite large
and does not actually equal the specified domains of the query or output construction
methods. This mismatch is common -- in fact, nearly universal. Therefore, 
our implementation includes a simple subtyping relation between regular expression
types. 
This subtyping relation is justified theoretically by the fact that language inclusion is
decidable; see \cite{fulton12} for a formal definition of the subtyping relation.
Additionally, our extension remains composable because subtyping is defined on a type-by-type basis;
see \cite{fulton13} for a discussion of subtyping in Atlang (referred to there as Ace).

@@

The result of applying sanitize to input is appended to two functions which
construct a safe query and safe output.
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
XSS attacks.

Note that ${\sf input}$ does not actually meet these specifications without additional machinery.
The type of ${\sf input}$ is quite large
and does not actually equal the specified domains of the query or output construction
methods. This mismatch is common -- in fact, nearly universal. Therefore, 
our implementation includes a simple subtyping relation between regular expression
types. 
This subtyping relation is justified theoretically by the fact that language inclusion is
decidable; see \cite{fulton12} for a formal definition of the subtyping relation.
Additionally, our extension remains composable because subtyping is defined on a type-by-type basis;
see \cite{fulton13} for a discussion of subtyping in Atlang (referred to there as Ace).

>>>>

@@

The result of applying sanitize to input is appended in two functions which
construct a safe query (avoiding command injection) and safe HTML output (avoiding XSS-style attacks).
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
Cross-Site Scripting (XSS) attacks.

Note that ${\sf input}$ does not actually meet these specifications without additional machinery.
The type of ${\sf input}$ is quite large
and does not actually equal the specified domains of the query or output construction
methods. This mismatch is common -- in fact, nearly universal. Therefore, 
our implementation includes a simple subtyping relation between regular expression
types.

This subtyping relation is justified theoretically by the fact that language inclusion is
decidable; see \cite{fulton12} for a formal definition of the subtyping relation.
Additionally, our extension remains composable because subtyping is defined on a type-by-type basis;
see \cite{fulton13} for a discussion of subtyping in Atlang (referred to there as Ace).

@@
